---
title: "Project Title"
subtitle: "A STAT 139 Final Project"
author: "Yuyue Wang, Xiangru Shu, Chengye Liu, Chia Chi (Michelle) Ho"
date: "Due December 13, 2017"

fontsize: 11pt
geometry: margin=1in
header-includes:
  - \usepackage{graphicx}
  - \usepackage{enumerate}
  - \usepackage{verbatim}
  - \usepackage{amsmath}
  - \usepackage{subfigure}
  - \usepackage{parskip}
  - \usepackage[notextcomp]{kpfonts}
  - \usepackage{geometry}
  - \usepackage[T1]{fontenc}
  - \usepackage{inconsolata}
  - \usepackage[dvipsnames]{xcolor}
  - \DeclareMathOperator*{\argmin}{argmin}
  
output:
  pdf_document:
    fig_width: 5
    fig_height: 3.5
  html_document:
    smart: false
---

## Abstract

## Introduction
- Obesity is an exerbating problem in the US. 
- Explore the association of 21 different factors with bmi, 13 of which are behavior-related factors such as the typical number of hours sleep per night


## Methods
- Data description
   - data source is NHANES 2013-2014
   - Variables of interest
   - Only consider adults of age 20 or above
- Data preprocessing & assumptions
  - Merge data by participant sequence number
  - Exclude don't know/refused/missing values --- discuss implications in limitations
- Perform EDA 
- Fit regression models
- Check assumptions

## Results
### Exploratory Data Analysis

## Limitations

## Conclusions 




## Appendix
### Appendix I: Data preprocessing
```{r data_preprocessing, warnings=FALSE, echo=FALSE}
library(dplyr)
demographics = read.csv("data/health/demographic.csv")
ques = read.csv("data/health/questionnaire.csv")
exam = read.csv("data/health/examination.csv")

# join tables by participant ID
df = inner_join(demographics, ques, by="SEQN")
df = inner_join(df, exam, by="SEQN")

# subset df to consider only the variables we're interested in
response = "BMXBMI"
predictors = c("RIAGENDR", "RIDAGEYR", "RIDRETH3", "DMDEDUC2", "DMDMARTL", "DMDFMSIZ", "INDFMIN2", 
               "ALQ101", "ALQ120Q", 
               "CBD070", "CBD120", "CBD130", 
               "DBD895", "DBD900", "DBQ197", 
               "DPQ020", "DPQ030",
               "PAQ710", "PAQ665",
               "SLD010H", "SMQ040")

columns = c(predictors, response)

df = df[names(df) %in% columns]

# rename columns to more intuitive names
df = rename(df, gender = RIAGENDR, age = RIDAGEYR, race = RIDRETH3, edu = DMDEDUC2,
            marriage = DMDMARTL, famsize = DMDFMSIZ, famincome = INDFMIN2,
            alcohol12yr = ALQ101, alcoholfrq = ALQ120Q, grocery = CBD070, eatout = CBD120,
            delivery = CBD130, milk = DBQ197, meals_nothome = DBD895, meals_fastfood = DBD900,
            depressed = DPQ020, sleep_trouble = DPQ030, tv_hrs = PAQ710,
            sleep_hr = SLD010H, smoke = SMQ040, bmi = BMXBMI, activity=PAQ665)

# subset the df to consider only adults aged 20 or above
df_adult = df[df$age > 20,]

# subset the data to exclude refused/don't know/missing data
# demographics variables
df_adult = df_adult[which(df_adult$edu!=7 & df_adult$edu!=9),]
df_adult = df_adult[which(df_adult$activity!=7 & df_adult$activity!=9),]
df_adult = df_adult[which(df_adult$marriage!=77 & df_adult$marriage!=99),]
df_adult = df_adult[which(df_adult$famincome!=77 & df_adult$famincome!=99),]

# alcohol use variables
df_adult = df_adult[which(df_adult$alcohol12yr!=7 & df_adult$alcohol12yr!=9),]
df_adult = df_adult[which(df_adult$alcoholfrq!=777 & df_adult$alcoholfrq!=999),]

# consumer behavior variables
df_adult = df_adult[which(df_adult$grocery!=777777 & df_adult$grocery!=999999),]
df_adult = df_adult[which(df_adult$eatout!=777777 & df_adult$eatout!=999999),]
df_adult = df_adult[which(df_adult$delivery!=777777 & df_adult$delivery!=999999),]

# diet behavior variables
df_adult = df_adult[which(df_adult$meals_nothome != 5555 & df_adult$meals_nothome != 7777 & df_adult$meals_nothome != 9999),]
df_adult = df_adult[which(df_adult$meals_fastfood != 5555 & df_adult$meals_fastfood != 7777 & df_adult$meals_fastfood != 9999),]

# physical activity variables
df_adult = df_adult[which(df_adult$tv_hrs != 77 & df_adult$tv_hrs != 99),]

# mental health variables
df_adult = df_adult[which(df_adult$depressed!=7 & df_adult$depressed!=9),]
df_adult = df_adult[which(df_adult$sleep_trouble!=7 & df_adult$sleep_trouble!=9),]

# sleeping behavior variables
df_adult = df_adult[which(df_adult$sleep_hr != 99),]

# smoking behavior variables
df_adult$smoke[which(is.na(df_adult$smoke))] = "missing"
df_adult$smoke[which(is.na(df_adult$activity))] = "missing"

# drop missing bmi observatioins
df_adult = df_adult[!is.na(df_adult$bmi),]


# feature engineering
df_adult$bmi_class = df_adult$bmi
df_adult$bmi_class[which(df_adult$bmi_class >= 40)] = "class 3 obesity"
df_adult$bmi_class[which(df_adult$bmi_class >= 30 & df_adult$bmi_class < 40)] = "obese"
df_adult$bmi_class[which(df_adult$bmi_class >= 25 & df_adult$bmi_class < 30)] = "overweight"
df_adult$bmi_class[which(df_adult$bmi_class >= 18.5 & df_adult$bmi_class < 25)] = "healthy"
df_adult$bmi_class[which(df_adult$bmi_class < 18.5)] = "underweight"

# save variable names into lists of categorical or numeric faetures
categorical_features = c("gender", "race", "edu", "marriage",
                         "famincome", "alcohol12yr", "milk", "depressed", 
                         "sleep_trouble", "smoke", "tv_hrs","activity")

numeric_features = c("height","age", "famsize", "alcoholfrq", "grocery", "eatout",
                     "delivery", "meals_nothome", "meals_fastfood", "sleep_hr")

#drop famincome 12,13
df_adult = df_adult[!(df_adult$famincome==12 | df_adult$famincome==13),]

# convert categorical variables into factors
df_adult[categorical_features] = lapply(df_adult[categorical_features], factor)

#activity
levels(df_adult$activity) = c(1, 2)
levels(df_adult$activity) = c("Yes", "No")

#gender
levels(df_adult$gender) = c("male", "female")

#race
levels(df_adult$race) = c("mexican", "other hispanic", "white", "black", "asian", "other")

#edu
levels(df_adult$edu) = c(1, 1, 2, 3, 3)
levels(df_adult$edu) = c("below hs", "hs", "above hs")

#alcohol consumption
levels(df_adult$alcohol12yr) = c("yes", "no")

#milk
levels(df_adult$milk) = c("never", "rare", "sometimes", "often", "varied")

# depressed
levels(df_adult$depressed) = c("no", "sometimes", "often", "alwyas")

# sleep_trouble
levels(df_adult$sleep_trouble) = c("no", "sometimes", "often", "always")

#smoke
levels(df_adult$smoke) = c("everyday", "sometimes", "no", "missing")
  
#famincome
levels(df_adult$famincome) = c(rep(1, 5), rep(2, 3), rep(3, 2), 4, 5 )
levels(df_adult$famincome) = c("below 25k", "25k~55k", "55k~75k", "75k~100k", "above 100k")

#marriage
levels(df_adult$marriage) = c("married", "widowed", "divorced", "divorced", "never_married", "married")

#tv_hrs
levels(df_adult$tv_hrs) = c("1hr","1hr", "2~4hrs", "2~4hrs" , "2~4hrs", "5hrs", "0hr")


apply(df_adult, 2, function(x) sum(is.na(x))) # check how many missing data
sapply(df_adult, class) # check data classes
```

### Appendix II: Exploratory Data Analysis
#### Response Variable (bmi)
```{r fig.height=3, fig.width=10, fig.align="center", warning=FALSE, echo=FALSE}
library(ggplot2)
require(gridExtra)
transform1 = log(df_adult$bmi)
transform2 = log(log(df_adult$bmi))

# response variable distribution
plot1 = ggplot(df_adult, aes(bmi)) +
  geom_histogram() +
  ggtitle("BMI")
plot2 = ggplot(df_adult, aes(transform1)) +
  geom_histogram() +
  ggtitle("log(bmi)")
plot3 = ggplot(df_adult, aes(transform2)) +
  geom_histogram() +
  ggtitle("Log(log(bmi))")
grid.arrange(plot1, plot2, plot3, ncol=3)
```


#### Predictor Variables



##### Numeric Variables
Distribtuion of numeric variables
```{r fig.height=10, fig.width=10, fig.align="center", warning=FALSE, echo=FALSE}
# numeric predictor variable distributions
plot1 = ggplot(df_adult, aes(age)) + geom_histogram() +
  ggtitle("Age")
plot3 = ggplot(df_adult, aes(famsize)) + geom_histogram() +
  ggtitle("Family Size")
plot4 = ggplot(df_adult, aes(alcoholfrq)) + geom_histogram() +
  ggtitle("Frequency of Alcohol Consumption")
plot5 = ggplot(df_adult, aes(grocery)) + geom_histogram() +
  ggtitle("Money Spent on Grocery")
plot6 = ggplot(df_adult, aes(eatout)) + geom_histogram() +
  ggtitle("Money Spent Eating Out")
plot7 = ggplot(df_adult, aes(delivery)) + geom_histogram() +
  ggtitle("Money Spent on Delivered Food")
plot8 = ggplot(df_adult, aes(meals_nothome)) + geom_histogram() +
  ggtitle("Number of Meals Not Home Prepared")
plot9 = ggplot(df_adult, aes(meals_fastfood)) + geom_histogram() +
  ggtitle("Number of Fast Food Meals")
plot10 = ggplot(df_adult, aes(sleep_hr)) + geom_histogram() +
  ggtitle("Hours of Sleep")
grid.arrange(plot1, plot3, plot4, plot5, plot6, plot7, plot8, plot9, plot10, ncol=3)
```                    
          
```{r fig.height=10, fig.width=9, fig.align="center", warning=FALSE, echo=FALSE}
# transformations on numeric predictor variables
plot1 = ggplot(df_adult, aes(alcoholfrq)) + geom_histogram() +
  ggtitle("Frequency of Alcohol Consumption")
plot2 = ggplot(df_adult, aes(log(alcoholfrq+1))) + geom_histogram() +
  ggtitle("Log Frequency of Alcohol Consumption")

plot3 = ggplot(df_adult, aes(grocery)) + geom_histogram() +
  ggtitle("Money Spent on Grocery")
plot4 = ggplot(df_adult, aes(log(grocery+1))) + geom_histogram() +
  ggtitle("Log Money Spent on Grocery")

plot5 = ggplot(df_adult, aes(eatout)) + geom_histogram() +
  ggtitle("Money Spent on Eating Out")
plot6 = ggplot(df_adult, aes(log(eatout+1))) + geom_histogram() +
  ggtitle("Log Money Spent on Eating Out")

plot7 = ggplot(df_adult, aes(delivery)) + geom_histogram() +
  ggtitle("Money Spent on Delivered Food")
plot8 = ggplot(df_adult, aes(log(delivery+1))) + geom_histogram() +
  ggtitle("Log Money Spent on Delivered Food")

plot9 = ggplot(df_adult, aes(meals_nothome)) + geom_histogram() +
  ggtitle("Number of Meals Not Home Prepared")
plot10 = ggplot(df_adult, aes(log(meals_nothome+1))) + geom_histogram() +
  ggtitle("Log Number of Meals Not Home Prepared")

plot11 = ggplot(df_adult, aes(meals_fastfood)) + geom_histogram() +
  ggtitle("Number of Fast Food Meals")
plot12 = ggplot(df_adult, aes(log(meals_fastfood+1))) + geom_histogram() +
  ggtitle("Log Number of Fast Food Meals")
grid.arrange(plot1, plot2, plot3, plot4, plot5, plot6, plot7, plot8, plot9, plot10, plot11, plot12, ncol=3)
```  


```{r data-transformation, warning=FALSE, echo=FALSE}
# response
#df_adult$log.bmi = log(df_adult$bmi)
df_adult$log2.bmi = log(log(df_adult$bmi))

# numeric predictors
#df_adult$log.alcoholfrq = log(df_adult$alcoholfrq + 1)
df_adult$log.grocery = log(df_adult$grocery + 1)
df_adult$log.eatout = log(df_adult$eatout + 1)
df_adult$log.delivery = log(df_adult$delivery + 1)
#df_adult$log.nothome = log(df_adult$meals_nothome + 1)
#df_adult$log.fastfood = log(df_adult$meals_fastfood + 1)
```


##### Categorical Variables
```{r fig.height=20, fig.width=15, fig.align="center", warning=FALSE, echo=FALSE}
plot1 = ggplot(df_adult, aes(x=gender, y = log2.bmi)) + geom_boxplot() + 
  theme(axis.text.x = element_text(angle=45),axis.text=element_text(size=16),title =element_text(size=18, face='bold'),axis.title=element_text(size=16,face="bold")) + 
  labs(x="Gender", y="log2.bmi") + 
  ggtitle("bmi by Gender") 
plot2 = ggplot(df_adult, aes(x=race, y = log2.bmi)) + geom_boxplot() + 
  theme(axis.text.x = element_text(angle=45),axis.text=element_text(size=16),title =element_text(size=18, face='bold'),axis.title=element_text(size=16,face="bold")) + 
  labs(x="Race", y="log2.bmi") + 
  ggtitle("bmi by Race")
plot3 = ggplot(df_adult, aes(x=edu, y = log2.bmi)) + geom_boxplot() + 
  theme(axis.text.x = element_text(angle=45),axis.text=element_text(size=16),title =element_text(size=18, face='bold'),axis.title=element_text(size=16,face="bold")) + 
  labs(x="Education Level", y="log2.bmi") + 
  ggtitle("bmi by Education Level") 
plot4 = ggplot(df_adult, aes(x=marriage, y = log2.bmi)) + geom_boxplot() + 
  theme(axis.text.x = element_text(angle=45),axis.text=element_text(size=16),title =element_text(size=18, face='bold'),axis.title=element_text(size=16,face="bold")) + 
  labs(x="Marital Status", y="log2.bmi") + 
  ggtitle("bmi by Marital Status") 
plot5 = ggplot(df_adult, aes(x=famincome, y = log2.bmi)) + geom_boxplot() + 
  theme(axis.text.x = element_text(angle=45),axis.text=element_text(size=16),title =element_text(size=18, face='bold'),axis.title=element_text(size=16,face="bold")) + 
  labs(x="Family Income", y="log2.bmi") + 
  ggtitle("bmi by Family Income")
plot6 = ggplot(df_adult, aes(x=alcohol12yr, y = log2.bmi)) + geom_boxplot() + 
  theme(axis.text.x = element_text(angle=45),axis.text=element_text(size=16),title =element_text(size=18, face='bold'),axis.title=element_text(size=16,face="bold")) + 
  labs(x="Alcohol Use", y="log2.bmi") + 
  ggtitle("bmi by Alcohol Use") 
plot7 = ggplot(df_adult, aes(x=milk, y = log2.bmi)) + geom_boxplot() + 
  theme(axis.text.x = element_text(angle=45),axis.text=element_text(size=16),title =element_text(size=18, face='bold'),axis.title=element_text(size=16,face="bold")) + 
  labs(x="Milk Consumption", y="log2.bmi") + 
  ggtitle("bmi by Milk Consumption") 
plot8 = ggplot(df_adult, aes(x=depressed, y = log2.bmi)) + geom_boxplot() + 
  theme(axis.text.x = element_text(angle=45),axis.text=element_text(size=16),title =element_text(size=18, face='bold'),axis.title=element_text(size=16,face="bold")) + 
  labs(x="Depressed", y="log2.bmi") + 
  ggtitle("bmi by Despressed") 
plot9 = ggplot(df_adult, aes(x=sleep_trouble, y = log2.bmi)) + geom_boxplot() + 
  theme(axis.text.x = element_text(angle=45),axis.text=element_text(size=16),title =element_text(size=18, face='bold'),axis.title=element_text(size=16,face="bold")) + 
  labs(x="Alcohol Use", y="log2.bmi") + 
  ggtitle("bmi by Sleep Trouble") 
plot10 = ggplot(df_adult, aes(x=smoke, y = log2.bmi)) + geom_boxplot() + 
  theme(axis.text.x = element_text(angle=45),axis.text=element_text(size=16),title =element_text(size=18, face='bold'),axis.title=element_text(size=16,face="bold")) + 
  labs(x="Smoke", y="log2.bmi") + 
  ggtitle("bmi by Smoke") 
plot11 = ggplot(df_adult, aes(x=tv_hrs, y = log2.bmi)) + geom_boxplot() + 
  theme(axis.text.x = element_text(angle=45),axis.text=element_text(size=16),title =element_text(size=18, face='bold'),axis.title=element_text(size=16,face="bold")) + 
  labs(x="TV Consumption", y="log2.bmi") + 
  ggtitle("bmi by TV Consumption")
plot12 = ggplot(df_adult, aes(x=activity, y = log2.bmi)) + geom_boxplot() + 
  theme(axis.text.x = element_text(angle=45),axis.text=element_text(size=16),title =element_text(size=18, face='bold'),axis.title=element_text(size=16,face="bold")) + 
  labs(x="Moderate Activities At Least 10 min Continuously per Week", y="log2.bmi") + 
  ggtitle("bmi by Moderate Recreational Activities") 
grid.arrange(plot1, plot2, plot3, plot4, plot5, plot6, plot7, plot8, plot9, plot10, plot11, plot12, ncol=3)
```

Response vs. numeric distribution
```{r fig.height=10, fig.width=10, fig.align="center", warning=FALSE, echo=FALSE}
plot1 = ggplot(df_adult, aes(x=age, y=log2.bmi)) + geom_point() + 
  ggtitle("Age")
plot3 = ggplot(df_adult, aes(x=famsize, y=log2.bmi)) + geom_point() + 
  ggtitle("Family Size")
plot4 = ggplot(df_adult, aes(x=alcoholfrq, y=log2.bmi)) + geom_point() + 
  ggtitle("Frequency of Alcohol Consumption")
plot5 = ggplot(df_adult, aes(x=log.grocery, y=log2.bmi)) + geom_point() + 
  ggtitle("Money Spent on Grocery")
plot6 =ggplot(df_adult, aes(x=log.eatout, y=log2.bmi)) + geom_point() + 
  ggtitle("Money Spent Eating Out")
plot7 = ggplot(df_adult, aes(x=log.delivery, y=log2.bmi)) + geom_point() + 
  ggtitle("Money Spent on Delivered Food")
plot8 =ggplot(df_adult, aes(x=meals_nothome, y=log2.bmi)) + geom_point() + 
  ggtitle("Number of Meals Not Home Prepared")
plot9 = ggplot(df_adult, aes(x=meals_fastfood, y=log2.bmi)) + geom_point() + 
  ggtitle("Number of Fast Food Meals")
plot10 = ggplot(df_adult, aes(x=sleep_hr, y=log2.bmi)) + geom_point() + 
  ggtitle("Hours of Sleep")
grid.arrange(plot1, plot3, plot4, plot5, plot6, plot7, plot8, plot9, plot10, ncol=3)
```          









