---
title: "Project Title"
subtitle: "A STAT 139 Final Project"
author: "Yuyue Wang, Xiangru Shu, Chengye Liu, Chia Chi (Michelle) Ho"
date: "Due December 13, 2017"

fontsize: 11pt
geometry: margin=1in
header-includes:
  - \usepackage{graphicx}
  - \usepackage{enumerate}
  - \usepackage{verbatim}
  - \usepackage{amsmath}
  - \usepackage{subfigure}
  - \usepackage{parskip}
  - \usepackage[notextcomp]{kpfonts}
  - \usepackage{geometry}
  - \usepackage[T1]{fontenc}
  - \usepackage{inconsolata}
  - \usepackage[dvipsnames]{xcolor}
  - \DeclareMathOperator*{\argmin}{argmin}
  
output:
  pdf_document:
    fig_width: 5
    fig_height: 3.5
  html_document:
    smart: false
---

## Abstract

## Introduction
- Obesity is an exerbating problem in the US. 
- Explore the association of 21 different factors with bmi, 13 of which are behavior-related factors such as the typical number of hours sleep per night


## Methods
- Data description
   - data source is NHANES 2013-2014
   - Variables of interest
   - Only consider adults of age 20 or above
- Data preprocessing & assumptions
  - Merge data by participant sequence number
  - Exclude don't know/refused/missing values --- discuss implications in limitations
- Perform EDA 
- Fit regression models
- Check assumptions

## Results
### Exploratory Data Analysis

## Limitations

## Conclusions 




## Appendix
### Appendix I: Data preprocessing
```{r data_preprocessing, warnings=FALSE, echo=FALSE}

set.seed(1234)
library(dplyr)
demographics = read.csv("data/health/demographic.csv")
ques = read.csv("data/health/questionnaire.csv")
exam = read.csv("data/health/examination.csv")

# join tables by participant ID
df = inner_join(demographics, ques, by="SEQN")
df = inner_join(df, exam, by="SEQN")

# subset df to consider only the variables we're interested in
response = "BMXBMI"
predictors = c("RIAGENDR", "RIDAGEYR", "RIDRETH3", "DMDEDUC2", "DMDMARTL", "DMDFMSIZ", "INDFMIN2", 
               "ALQ101", "ALQ120Q", 
               "CBD070", "CBD120", "CBD130", 
               "DBD895", "DBD900", "DBQ197", 
               "DPQ020", "DPQ030",
               "PAQ710", "PAQ665",
               "SLD010H", "SMQ040")

columns = c(predictors, response)

df = df[names(df) %in% columns]

# rename columns to more intuitive names
df = rename(df, gender = RIAGENDR, age = RIDAGEYR, race = RIDRETH3, edu = DMDEDUC2,
            marriage = DMDMARTL, famsize = DMDFMSIZ, famincome = INDFMIN2,
            alcohol12yr = ALQ101, alcoholfrq = ALQ120Q, grocery = CBD070, eatout = CBD120,
            delivery = CBD130, milk = DBQ197, meals_nothome = DBD895, meals_fastfood = DBD900,
            depressed = DPQ020, sleep_trouble = DPQ030, tv_hrs = PAQ710,
            sleep_hr = SLD010H, smoke = SMQ040, bmi = BMXBMI, activity=PAQ665)

# subset the df to consider only adults aged 20 or above
df_adult = df[df$age > 21,]

# subset the data to exclude refused/don't know/missing data
# demographics variables
df_adult = df_adult[which(df_adult$edu!=7 & df_adult$edu!=9),]
df_adult = df_adult[which(df_adult$activity!=7 & df_adult$activity!=9),]
df_adult = df_adult[which(df_adult$marriage!=77 & df_adult$marriage!=99),]
df_adult = df_adult[which(df_adult$famincome!=77 & df_adult$famincome!=99),]

# alcohol use variables
df_adult = df_adult[which(df_adult$alcohol12yr!=7 & df_adult$alcohol12yr!=9),]
df_adult = df_adult[which(df_adult$alcoholfrq!=777 & df_adult$alcoholfrq!=999),]

# consumer behavior variables
df_adult = df_adult[which(df_adult$grocery!=777777 & df_adult$grocery!=999999),]
df_adult = df_adult[which(df_adult$eatout!=777777 & df_adult$eatout!=999999),]
df_adult = df_adult[which(df_adult$delivery!=777777 & df_adult$delivery!=999999),]

# diet behavior variables
df_adult = df_adult[which(df_adult$meals_nothome != 5555 & df_adult$meals_nothome != 7777 & df_adult$meals_nothome != 9999),]
df_adult = df_adult[which(df_adult$meals_fastfood != 5555 & df_adult$meals_fastfood != 7777 & df_adult$meals_fastfood != 9999),]

# physical activity variables
df_adult = df_adult[which(df_adult$tv_hrs != 77 & df_adult$tv_hrs != 99),]

# mental health variables
df_adult = df_adult[which(df_adult$depressed!=7 & df_adult$depressed!=9),]
df_adult = df_adult[which(df_adult$sleep_trouble!=7 & df_adult$sleep_trouble!=9),]

# sleeping behavior variables
df_adult = df_adult[which(df_adult$sleep_hr != 99),]

# smoking behavior variables
df_adult$smoke[which(is.na(df_adult$smoke))] = "missing"
df_adult$smoke[which(is.na(df_adult$activity))] = "missing"

# drop missing bmi observatioins
df_adult = df_adult[!is.na(df_adult$bmi),]


# feature engineering
df_adult$bmi_class = df_adult$bmi
# df_adult$bmi_class[which(df_adult$bmi_class >= 40)] = "class 3 obesity"
df_adult$bmi_class[which(df_adult$bmi_class >= 30 )] = "obese"
df_adult$bmi_class[which(df_adult$bmi_class >= 25 & df_adult$bmi_class < 30)] = "overweight"
df_adult$bmi_class[which(df_adult$bmi_class >= 18.5 & df_adult$bmi_class < 25)] = "healthy"
df_adult$bmi_class[which(df_adult$bmi_class < 18.5)] = "underweight"

# save variable names into lists of categorical or numeric faetures
categorical_features = c("gender", "race", "edu", "marriage",
                         "famincome", "alcohol12yr", "milk", "depressed", 
                         "sleep_trouble", "smoke", "tv_hrs","activity")

numeric_features = c("height","age", "famsize", "alcoholfrq", "grocery", "eatout",
                     "delivery", "meals_nothome", "meals_fastfood", "sleep_hr")

#drop famincome 12,13
df_adult = df_adult[!(df_adult$famincome==12 | df_adult$famincome==13),]

# convert categorical variables into factors
df_adult[categorical_features] = lapply(df_adult[categorical_features], factor)

#activity
levels(df_adult$activity) = c(1, 2)
levels(df_adult$activity) = c("Yes", "No")

#gender
levels(df_adult$gender) = c("male", "female")

#race
levels(df_adult$race) = c("mexican", "other hispanic", "white", "black", "asian", "other")

#edu
levels(df_adult$edu) = c(1, 1, 2, 3, 3)
levels(df_adult$edu) = c("below hs", "hs", "above hs")

#alcohol consumption
levels(df_adult$alcohol12yr) = c("yes", "no")

#milk
levels(df_adult$milk) = c("never", "rare", "sometimes", "often", "varied")

# depressed
levels(df_adult$depressed) = c("no", "sometimes", "often", "alwyas")

# sleep_trouble
levels(df_adult$sleep_trouble) = c("no", "sometimes", "often", "always")

#smoke
levels(df_adult$smoke) = c("everyday", "sometimes", "no", "missing")
  
#famincome
levels(df_adult$famincome) = c(rep(1, 5), rep(2, 3), rep(3, 2), 4, 5 )
levels(df_adult$famincome) = c("below 25k", "25k~55k", "55k~75k", "75k~100k", "above 100k")

#marriage
levels(df_adult$marriage) = c("married", "widowed", "divorced", "divorced", "never_married", "married")

#tv_hrs
levels(df_adult$tv_hrs) = c("1hr","1hr", "2~4hrs", "2~4hrs" , "2~4hrs", "5hrs", "0hr")


apply(df_adult, 2, function(x) sum(is.na(x))) # check how many missing data
sapply(df_adult, class) # check data classes
```

### Appendix II: Exploratory Data Analysis
#### Response Variable (bmi)
```{r fig.height=3, fig.width=10, fig.align="center", warning=FALSE, echo=FALSE}
library(ggplot2)
require(gridExtra)
transform1 = log(df_adult$bmi)
transform2 = log(log(df_adult$bmi))

# response variable distribution
plot1 = ggplot(df_adult, aes(bmi)) +
  geom_histogram() +
  ggtitle("BMI")
plot2 = ggplot(df_adult, aes(transform1)) +
  geom_histogram() +
  ggtitle("log(bmi)")
plot3 = ggplot(df_adult, aes(transform2)) +
  geom_histogram() +
  ggtitle("Log(log(bmi))")
grid.arrange(plot1, plot2, plot3, ncol=3)
```


#### Predictor Variables



##### Numeric Variables
Distribtuion of numeric variables
```{r fig.height=10, fig.width=10, fig.align="center", warning=FALSE, echo=FALSE}
# numeric predictor variable distributions
plot1 = ggplot(df_adult, aes(age)) + geom_histogram() +
  ggtitle("Age")
plot3 = ggplot(df_adult, aes(famsize)) + geom_histogram() +
  ggtitle("Family Size")
plot4 = ggplot(df_adult, aes(alcoholfrq)) + geom_histogram() +
  ggtitle("Frequency of Alcohol Consumption")
plot5 = ggplot(df_adult, aes(grocery)) + geom_histogram() +
  ggtitle("Money Spent on Grocery")
plot6 = ggplot(df_adult, aes(eatout)) + geom_histogram() +
  ggtitle("Money Spent Eating Out")
plot7 = ggplot(df_adult, aes(delivery)) + geom_histogram() +
  ggtitle("Money Spent on Delivered Food")
plot8 = ggplot(df_adult, aes(meals_nothome)) + geom_histogram() +
  ggtitle("Number of Meals Not Home Prepared")
plot9 = ggplot(df_adult, aes(meals_fastfood)) + geom_histogram() +
  ggtitle("Number of Fast Food Meals")
plot10 = ggplot(df_adult, aes(sleep_hr)) + geom_histogram() +
  ggtitle("Hours of Sleep")
grid.arrange(plot1, plot3, plot4, plot5, plot6, plot7, plot8, plot9, plot10, ncol=3)
```                    
          
```{r fig.height=10, fig.width=9, fig.align="center", warning=FALSE, echo=FALSE}
# transformations on numeric predictor variables
plot1 = ggplot(df_adult, aes(alcoholfrq)) + geom_histogram() +
  ggtitle("Frequency of Alcohol Consumption")
plot2 = ggplot(df_adult, aes(log(alcoholfrq+1))) + geom_histogram() +
  ggtitle("Log Frequency of Alcohol Consumption")

plot3 = ggplot(df_adult, aes(grocery)) + geom_histogram() +
  ggtitle("Money Spent on Grocery")
plot4 = ggplot(df_adult, aes(log(grocery+1))) + geom_histogram() +
  ggtitle("Log Money Spent on Grocery")

plot5 = ggplot(df_adult, aes(eatout)) + geom_histogram() +
  ggtitle("Money Spent on Eating Out")
plot6 = ggplot(df_adult, aes(log(eatout+1))) + geom_histogram() +
  ggtitle("Log Money Spent on Eating Out")

plot7 = ggplot(df_adult, aes(delivery)) + geom_histogram() +
  ggtitle("Money Spent on Delivered Food")
plot8 = ggplot(df_adult, aes(log(delivery+1))) + geom_histogram() +
  ggtitle("Log Money Spent on Delivered Food")

plot9 = ggplot(df_adult, aes(meals_nothome)) + geom_histogram() +
  ggtitle("Number of Meals Not Home Prepared")
plot10 = ggplot(df_adult, aes(log(meals_nothome+1))) + geom_histogram() +
  ggtitle("Log Number of Meals Not Home Prepared")

plot11 = ggplot(df_adult, aes(meals_fastfood)) + geom_histogram() +
  ggtitle("Number of Fast Food Meals")
plot12 = ggplot(df_adult, aes(log(meals_fastfood+1))) + geom_histogram() +
  ggtitle("Log Number of Fast Food Meals")
grid.arrange(plot1, plot2, plot3, plot4, plot5, plot6, plot7, plot8, plot9, plot10, plot11, plot12, ncol=3)
```  


```{r data-transformation, warning=FALSE, echo=FALSE}
# response
#df_adult$log.bmi = log(df_adult$bmi)
df_adult$log2.bmi = log(log(df_adult$bmi))

# numeric predictors
#df_adult$log.alcoholfrq = log(df_adult$alcoholfrq + 1)
df_adult$log.grocery = log(df_adult$grocery + 1)
df_adult$log.eatout = log(df_adult$eatout + 1)
df_adult$log.delivery = log(df_adult$delivery + 1)
#df_adult$log.nothome = log(df_adult$meals_nothome + 1)
#df_adult$log.fastfood = log(df_adult$meals_fastfood + 1)
```


##### Categorical Variables
```{r fig.height=20, fig.width=15, fig.align="center", warning=FALSE, echo=FALSE}
plot1 = ggplot(df_adult, aes(x=gender, y = log2.bmi)) + geom_boxplot() + 
  theme(axis.text.x = element_text(angle=45),axis.text=element_text(size=16),title =element_text(size=18, face='bold'),axis.title=element_text(size=16,face="bold")) + 
  labs(x="Gender", y="log2.bmi") + 
  ggtitle("bmi by Gender") 
plot2 = ggplot(df_adult, aes(x=race, y = log2.bmi)) + geom_boxplot() + 
  theme(axis.text.x = element_text(angle=45),axis.text=element_text(size=16),title =element_text(size=18, face='bold'),axis.title=element_text(size=16,face="bold")) + 
  labs(x="Race", y="log2.bmi") + 
  ggtitle("bmi by Race")
plot3 = ggplot(df_adult, aes(x=edu, y = log2.bmi)) + geom_boxplot() + 
  theme(axis.text.x = element_text(angle=45),axis.text=element_text(size=16),title =element_text(size=18, face='bold'),axis.title=element_text(size=16,face="bold")) + 
  labs(x="Education Level", y="log2.bmi") + 
  ggtitle("bmi by Education Level") 
plot4 = ggplot(df_adult, aes(x=marriage, y = log2.bmi)) + geom_boxplot() + 
  theme(axis.text.x = element_text(angle=45),axis.text=element_text(size=16),title =element_text(size=18, face='bold'),axis.title=element_text(size=16,face="bold")) + 
  labs(x="Marital Status", y="log2.bmi") + 
  ggtitle("bmi by Marital Status") 
plot5 = ggplot(df_adult, aes(x=famincome, y = log2.bmi)) + geom_boxplot() + 
  theme(axis.text.x = element_text(angle=45),axis.text=element_text(size=16),title =element_text(size=18, face='bold'),axis.title=element_text(size=16,face="bold")) + 
  labs(x="Family Income", y="log2.bmi") + 
  ggtitle("bmi by Family Income")
plot6 = ggplot(df_adult, aes(x=alcohol12yr, y = log2.bmi)) + geom_boxplot() + 
  theme(axis.text.x = element_text(angle=45),axis.text=element_text(size=16),title =element_text(size=18, face='bold'),axis.title=element_text(size=16,face="bold")) + 
  labs(x="Alcohol Use", y="log2.bmi") + 
  ggtitle("bmi by Alcohol Use") 
plot7 = ggplot(df_adult, aes(x=milk, y = log2.bmi)) + geom_boxplot() + 
  theme(axis.text.x = element_text(angle=45),axis.text=element_text(size=16),title =element_text(size=18, face='bold'),axis.title=element_text(size=16,face="bold")) + 
  labs(x="Milk Consumption", y="log2.bmi") + 
  ggtitle("bmi by Milk Consumption") 
plot8 = ggplot(df_adult, aes(x=depressed, y = log2.bmi)) + geom_boxplot() + 
  theme(axis.text.x = element_text(angle=45),axis.text=element_text(size=16),title =element_text(size=18, face='bold'),axis.title=element_text(size=16,face="bold")) + 
  labs(x="Depressed", y="log2.bmi") + 
  ggtitle("bmi by Despressed") 
plot9 = ggplot(df_adult, aes(x=sleep_trouble, y = log2.bmi)) + geom_boxplot() + 
  theme(axis.text.x = element_text(angle=45),axis.text=element_text(size=16),title =element_text(size=18, face='bold'),axis.title=element_text(size=16,face="bold")) + 
  labs(x="Alcohol Use", y="log2.bmi") + 
  ggtitle("bmi by Sleep Trouble") 
plot10 = ggplot(df_adult, aes(x=smoke, y = log2.bmi)) + geom_boxplot() + 
  theme(axis.text.x = element_text(angle=45),axis.text=element_text(size=16),title =element_text(size=18, face='bold'),axis.title=element_text(size=16,face="bold")) + 
  labs(x="Smoke", y="log2.bmi") + 
  ggtitle("bmi by Smoke") 
plot11 = ggplot(df_adult, aes(x=tv_hrs, y = log2.bmi)) + geom_boxplot() + 
  theme(axis.text.x = element_text(angle=45),axis.text=element_text(size=16),title =element_text(size=18, face='bold'),axis.title=element_text(size=16,face="bold")) + 
  labs(x="TV Consumption", y="log2.bmi") + 
  ggtitle("bmi by TV Consumption")
plot12 = ggplot(df_adult, aes(x=activity, y = log2.bmi)) + geom_boxplot() + 
  theme(axis.text.x = element_text(angle=45),axis.text=element_text(size=16),title =element_text(size=18, face='bold'),axis.title=element_text(size=16,face="bold")) + 
  labs(x="Moderate Activities At Least 10 min Continuously per Week", y="log2.bmi") + 
  ggtitle("bmi by Moderate Recreational Activities") 
grid.arrange(plot1, plot2, plot3, plot4, plot5, plot6, plot7, plot8, plot9, plot10, plot11, plot12, ncol=3)
```

Response vs. numeric distribution
```{r fig.height=10, fig.width=10, fig.align="center", warning=FALSE, echo=FALSE}
plot1 = ggplot(df_adult, aes(x=age, y=log2.bmi)) + geom_point() + 
  ggtitle("Age")
plot3 = ggplot(df_adult, aes(x=famsize, y=log2.bmi)) + geom_point() + 
  ggtitle("Family Size")
plot4 = ggplot(df_adult, aes(x=alcoholfrq, y=log2.bmi)) + geom_point() + 
  ggtitle("Frequency of Alcohol Consumption")
plot5 = ggplot(df_adult, aes(x=log.grocery, y=log2.bmi)) + geom_point() + 
  ggtitle("Money Spent on Grocery")
plot6 =ggplot(df_adult, aes(x=log.eatout, y=log2.bmi)) + geom_point() + 
  ggtitle("Money Spent Eating Out")
plot7 = ggplot(df_adult, aes(x=log.delivery, y=log2.bmi)) + geom_point() + 
  ggtitle("Money Spent on Delivered Food")
plot8 =ggplot(df_adult, aes(x=meals_nothome, y=log2.bmi)) + geom_point() + 
  ggtitle("Number of Meals Not Home Prepared")
plot9 = ggplot(df_adult, aes(x=meals_fastfood, y=log2.bmi)) + geom_point() + 
  ggtitle("Number of Fast Food Meals")
plot10 = ggplot(df_adult, aes(x=sleep_hr, y=log2.bmi)) + geom_point() + 
  ggtitle("Hours of Sleep")
grid.arrange(plot1, plot3, plot4, plot5, plot6, plot7, plot8, plot9, plot10, ncol=3)
```          



### Models
```{r}
#df_adult
check_assumptions <- function(model){
  hist(resid(model),col="grey",breaks=15)
  plot(resid(model)~fitted(model),cex=1.5)
  abline(h=0,lwd=2)
}
```





```{r fig.height=3, fig.width=3, fig.align="center"}
library(Matrix)
library(glmnet)

library(plotmo)
library(caret)
library(lattice)

set.seed(1234)
# define training control
train_control <- trainControl(method="cv", number=5)


#baseline model
model1 = lm(log2.bmi~gender+age+race+edu+marriage+famsize+famincome+alcohol12yr+alcoholfrq+log.grocery+log.eatout+log.delivery+milk+meals_nothome+meals_fastfood+depressed+sleep_trouble+activity+tv_hrs+sleep_hr+smoke, data = df_adult)

base_form = formula(model1)

summary(model1)
plotres(model1, which=3:4, caption = "Base model assumption check")

#baseline model cv
model1_cv = train(base_form, data=df_adult, method="lm", preProcess="scale", trControl=train_control)

print("Base model cv results:")
print(paste("RMSE:", model1_cv$results$RMSE, "Rsquared:", model1_cv$results$Rsquared))

```



```{r}
#model1a: baseline model step
set.seed(1234)
model1a = step(model1, trace = 0, direction = "backward")

summary(model1a)
plotres(model1a, which=3:4, caption = "model1a assumption check")

base_1a_form = formula(model1a)
#baseline model cv
model1a_cv = train(base_1a_form, data=df_adult, method="lm", preProcess="scale", trControl=train_control)

print("model1a cv results:")
print(paste("RMSE:", model1a_cv$results$RMSE, "Rsquared:", model1a_cv$results$Rsquared))
```


```{r}
#compare model1 with model1a
set.seed(1234)
anova(model1a, model1)
```



```{r fig.height=3, fig.width=3, fig.align="center"}
#model1b: add sleep^2 to model1a
set.seed(1234)

base_sleep_form = update.formula(base_1a_form, . ~. + I(sleep_hr^2))

model1b = lm(base_sleep_form, data = df_adult)
summary(model1b)
plotres(model1b, which=3:4, caption = "model1a plus sleephour^2 assumption check")

model1b_form = formula(model1b)

#model1b cv
model1b_cv = train(base_sleep_form, data=df_adult, method="lm", preProcess="scale", trControl=train_control)


print("model1a plus sleephour^2 cv results:")
print(paste("RMSE:", model1b_cv$results$RMSE, "Rsquared:", model1b_cv$results$Rsquared))



```


```{r}
#compare model1b to model1a
set.seed(1234)
anova(model1a, model1b)
```

```{r fig.height=3, fig.width=3, fig.align="center"}
#model1c: add famsize interactions
set.seed(1234)
model1c = lm(update.formula(model1b_form, .~.+famsize*(famincome+log.grocery+log.eatout+log.delivery)), data = df_adult)

summary(model1c)
extractAIC(model1c)
plotres(model1c, which=3:4, caption = "model1c assumption check")


#model1c cv
model1c_cv = train(formula(model1c), data=df_adult, method="lm", preProcess="scale", trControl=train_control)

print("Model1c cv results:")
print(paste("RMSE:", model1c_cv$results$RMSE, "Rsquared:", model1c_cv$results$Rsquared))


```
```{r}
#compare model1c to model1b
set.seed(1234)
anova(model1b, model1c)
```



```{r fig.height=3, fig.width=3, fig.align="center", warning=FALSE, echo=FALSE}
#model1d: the model with all the 2 way interactions(including sleep_hr2) after step()
set.seed(1234)
full_intr_form = update(model1b_form, .~ (.)^2)
model_interact = lm(full_intr_form, data = df_adult)

model1d = step(lm(log2.bmi ~ 1, data = df_adult), scope = list(upper = model_interact), direction = "forward", trace = 0)
summary(model1d)
plotres(model1d, which=3:4, caption = "model1d assumption check")

#model1d cv
model1d_cv = train(formula(model1d), data=df_adult, method="lm", preProcess="scale", trControl=train_control)

print("model1d cv results:")
print(paste("RMSE:", model1d_cv$results$RMSE, "Rsquared:", model1d_cv$results$Rsquared))

```


```{r}
#compare model1d to model1b
set.seed(1234)
anova(model1b, model1d)
```




```{r, fig.height=3, fig.width=3, fig.align="center"}

#lasso cv
set.seed(1234)
x.lasso = model.matrix(update(full_intr_form, .~.+0), data=df_adult)
cv_lasso = cv.glmnet(x.lasso, df_adult$log2.bmi, nfolds = 5)

lasso.lambda.min = cv_lasso$lambda.min
lasso.mse.min = min(cv_lasso$cvm)

coef.lasso = coef(cv_lasso, s = "lambda.min")
nonzero_coef_lasso = cbind(rownames(coef.lasso)[which(coef.lasso != 0)], coef.lasso[which(coef.lasso != 0)])

#lasso 
model_lasso = glmnet(x.lasso, df_adult$log2.bmi, lambda = lasso.lambda.min)

#plot
plot(cv_lasso, main="Lasso: MSE vs. lambda")
summary(model_lasso)
plotres(model_lasso, which=3:4, caption = "Full interaction Lasso model assumption check")
```
```{r}
lasso.mse.min
```



```{r, fig.height=3, fig.width=3, fig.align="center"}
#ridge cv
set.seed(1234)
x.ridge = model.matrix(update(full_intr_form, .~.+0), data=df_adult)
cv_ridge = cv.glmnet(x.ridge, df_adult$log2.bmi, alpha = 0, nfolds = 5)

ridge.lambda.min = cv_ridge$lambda.min
ridge.mse.min = min(cv_ridge$cvm)

coef.ridge = coef(cv_ridge, s = "lambda.min")
nonzero_coef_ridge = cbind(rownames(coef.ridge)[which(coef.ridge != 0)], coef.ridge[which(coef.ridge != 0)])

#ridge
model_ridge = glmnet(x.ridge, df_adult$log2.bmi, lambda = ridge.lambda.min, alpha = 0)

#plot
plot(cv_ridge, main="Ridge: MSE vs. lambda")
summary(model_ridge)
plotres(model_ridge, which=3:4, caption = "Full interaction Lasso model \n assumption check")
```
```{r}
ridge.mse.min
```




```{r, fig.height=3, fig.width=3, fig.align="center"}
#elastic net cv
set.seed(1234)
x.elastic = model.matrix(update(full_intr_form, .~.+0), data=df_adult)
cv_elastic = cv.glmnet(x.elastic, df_adult$log2.bmi, nfolds = 5)

elastic.lambda.min = cv_elastic$lambda.min
elastic.mse.min = min(cv_elastic$cvm)

coef.elastic = coef(cv_elastic, s = "lambda.min")
nonzero_coef_elastic = cbind(rownames(coef.elastic)[which(coef.elastic != 0)], coef.elastic[which(coef.elastic != 0)])

#elastic net 
model_elastic = glmnet(x.elastic, df_adult$log2.bmi, lambda = elastic.lambda.min, alpha=0.5)

#plot
plot(cv_elastic, main="Elastic net: MSE vs. lambda")
summary(model_elastic)
plotres(model_elastic, which=3:4, caption = "Full interaction Elastic model \n assumption check")

```
```{r}
elastic.mse.min
```


```{r}
# best model (model1d) bootstraping 

# Bootstrap 95% CI for regression coefficients 
set.seed(1234)
library(boot)
# function to obtain regression weights 
bs <- function(formula, data, indices) {
  d <- data[indices,] # allows boot to select sample 
  fit <- lm(formula, data=d)
  return(coef(fit)) 
} 
# bootstrapping with 1000 replications 
nsim = 500
results <- boot(data=df_adult, statistic=bs, R=nsim, formula=formula(model1d))
names_ = names(coef(lm(formula(model1d), data=df_adult)))

ci = data.frame(matrix(ncol = 3, nrow = length(names_)))

print("the following coef is statistically significant from 0:")
for (i in c(1:85, 88:90)){
  ci[i, 1:2] = boot.ci(boot.out=results, conf = 0.95, type = "perc", index= i)$perc[,4:5]
  ci[i, 3] = (0 < ci[i, 2]) & (0 > ci[i, 1]) 
  if (ci[i, 3] == FALSE) {print(names_[i])}
}

```

```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(nnet)
set.seed(1234)

#multinomial classification
## -- Base model
multinomial_base_form = update.formula(formula(base_form), bmi_class~.)
multinomial_base_cv = train(multinomial_base_form, data=df_adult, method="multinom", preProcess="scale", trControl=train_control, trace=0)

print(paste("Multinomial classification base model cv accuracy:", max(multinomial_base_cv$results$Accuracy)))

## -- Base sleep model
multinomial_base_sleep_form = update.formula(formula(base_sleep_form), bmi_class~.)
multinomial_base_sleep_cv = train(multinomial_base_sleep_form, data=df_adult, method="multinom", preProcess="scale", trControl=train_control, trace=0)

print(paste("Multinomial classification base sleep model cv accuracy:", max(multinomial_base_sleep_cv$results$Accuracy)))

## -- forward selected full interaction model
multinomial_1d =  update.formula(formula(model1d), bmi_class~.)

multinomial_1d_cv = train(multinomial_1d, data=df_adult, method="multinom", preProcess="scale", trControl=train_control, trace=0)

print(paste("Multinomial classification model1d cv results:", max(multinomial_1d_cv$results$Accuracy)))


multinomial_base_form = update.formula(formula(base_form), bmi_class~.)
multinomial_base_cv = train(multinomial_base_form, data=df_adult, method="multinom", preProcess="scale", trControl=train_control, trace=0)

print(paste("Ordinal classification base model cv accuracy:", max(multinomial_base_cv$results$Accuracy)))

library(ordinalNet)

#ordinal logistic regression
## -- Base model
x = model.matrix(update.formula(multinomial_base_form, .~.+0), data=df_adult)

ordinal_base_cv = ordinalNetCV(x, as.factor(df_adult$bmi_class), lambdaVals=c(0), nFolds=5)

print(paste("Ordinal classification base model cv accuracy:", 1-mean(summary(ordinal_base_cv)$misclass)))


```


```{r}
write.csv(data.frame(summary(model1)$coefficients), file="table1.csv")
write.csv(data.frame(summary(model1a)$coefficients), file="table1a.csv")
write.csv(data.frame(summary(model1b)$coefficients), file="table1b.csv")
write.csv(data.frame(summary(model1c)$coefficients), file="table1c.csv")

write.csv(data.frame(nonzero_coef_lasso), file="table_lasso.csv")
write.csv(data.frame(nonzero_coef_ridge), file="table_ridge.csv")
write.csv(data.frame(nonzero_coef_elastic), file="table_elastic.csv")
```

