---
title: "Project Title"
subtitle: "A STAT 139 Final Project"
author: "Yuyue Wang, Xiangru Shu, Chengye Liu, Chia Chi (Michelle) Ho"
date: "Due December 13, 2017"

fontsize: 11pt
geometry: margin=1in
header-includes:
  - \usepackage{graphicx}
  - \usepackage{enumerate}
  - \usepackage{verbatim}
  - \usepackage{amsmath}
  - \usepackage{subfigure}
  - \usepackage{parskip}
  - \usepackage[notextcomp]{kpfonts}
  - \usepackage{geometry}
  - \usepackage[T1]{fontenc}
  - \usepackage{inconsolata}
  - \usepackage[dvipsnames]{xcolor}
  - \DeclareMathOperator*{\argmin}{argmin}
  
output:
  pdf_document:
    fig_width: 5
    fig_height: 3.5
---

## Abstract

## Introduction
- Obesity is an exerbating problem in the US. 
- Explore the association of 21 different factors with weight, 13 of which are behavior-related factors such as the typical number of hours sleep per night


## Methods
- Data description
   - data source is NHANES 2013-2014
   - Variables of interest
   - Only consider adults of age 20 or above
- Data preprocessing & assumptions
  - Merge data by participant sequence number
  - Exclude don't know/refused/missing values --- discuss implications in limitations
- Perform EDA 
- Fit regression models
- Check assumptions

## Results
### Exploratory Data Analysis

## Limitations

## Conclusions 




## Appendix
### Appendix I: Data preprocessing
```{r data_preprocessing, warnings=FALSE}
library(dplyr)
demographics = read.csv("data/health/demographic.csv")
ques = read.csv("data/health/questionnaire.csv")
exam = read.csv("data/health/examination.csv")

# join tables by participant ID
df = inner_join(demographics, ques, by="SEQN")
df = inner_join(df, exam, by="SEQN")
dim(df)

# subset df to consider only the variables we're interested in
response = "BMXBMI"
predictors = c("RIAGENDR", "RIDAGEYR", "RIDRETH3", "DMDEDUC2", "DMDMARTL", "DMDFMSIZ", "INDFMIN2", 
               "ALQ101", "ALQ120Q", 
               "CBD070", "CBD120", "CBD130", 
               "DBD895", "DBD900", "DBQ197", 
               "DPQ020", "DPQ030",
               "PAQ710", 
               "SLD010H", "SMQ040")

columns = c(predictors, response)

df = df[names(df) %in% columns]

# rename columns to more intuitive names
df = rename(df, gender = RIAGENDR, age = RIDAGEYR, race = RIDRETH3, edu = DMDEDUC2,
            marriage = DMDMARTL, famsize = DMDFMSIZ, famincome = INDFMIN2,
            alcohol12yr = ALQ101, alcoholfrq = ALQ120Q, grocery = CBD070, eatout = CBD120,
            delivery = CBD130, milk = DBQ197, meals_nothome = DBD895, meals_fastfood = DBD900,
            depressed = DPQ020, sleep_trouble = DPQ030, tv_hrs = PAQ710,
            sleep_hr = SLD010H, smoke = SMQ040, bmi = BMXBMI)

# subset the df to consider only adults aged 20 or above
df_adult = df[df$age > 20,]

# subset the data to exclude refused/don't know/missing data
# demographics variables
df_adult = df_adult[which(df_adult$edu!=7 & df_adult$edu!=9),]
df_adult = df_adult[which(df_adult$marriage!=77 & df_adult$marriage!=99),]
df_adult = df_adult[which(df_adult$famincome!=77 & df_adult$famincome!=99),]

# alcohol use variables
df_adult = df_adult[which(df_adult$alcohol12yr!=7 & df_adult$alcohol12yr!=9),]
df_adult = df_adult[which(df_adult$alcoholfrq!=777 & df_adult$alcoholfrq!=999),]

# consumer behavior variables
df_adult = df_adult[which(df_adult$grocery!=777777 & df_adult$grocery!=999999),]
df_adult = df_adult[which(df_adult$eatout!=777777 & df_adult$eatout!=999999),]
df_adult = df_adult[which(df_adult$delivery!=777777 & df_adult$delivery!=999999),]

# diet behavior variables
df_adult = df_adult[which(df_adult$meals_nothome != 5555 & df_adult$meals_nothome != 7777 & df_adult$meals_nothome != 9999),]
df_adult = df_adult[which(df_adult$meals_fastfood != 5555 & df_adult$meals_fastfood != 7777 & df_adult$meals_fastfood != 9999),]

# physical activity variables
df_adult$tv_hrs[which(df_adult$tv_hrs == 0)] = 1
df_adult$tv_hrs[which(df_adult$tv_hrs == 8)] = 0
df_adult = df_adult[which(df_adult$tv_hrs != 77 & df_adult$tv_hrs != 99),]

# mental health variables
df_adult = df_adult[which(df_adult$depressed!=7 & df_adult$depressed!=9),]
df_adult = df_adult[which(df_adult$sleep_trouble!=7 & df_adult$sleep_trouble!=9),]

# sleeping behavior variables
df_adult = df_adult[which(df_adult$sleep_hr != 99),]

# smoking behavior variables
df_adult$smoke[which(is.na(df_adult$smoke))] = "missing"

# after dropping observations missing weight, there were only 3 missing height
# so we drop these observations too
drop_obs = c("weight", "height")

for (feature in drop_obs){
  df_adult = df_adult[!is.na(df_adult[feature]),]
}

# save variable names into lists of categorical or numeric faetures
categorical_features = c("gender", "race", "edu", "marriage",
                         "famincome", "alcohol12yr", "milk", "depressed", 
                         "sleep_trouble", "smoke", "tv_hrs")

numeric_features = c("height","age", "famsize", "alcoholfrq", "grocery", "eatout",
                     "delivery", "meals_nothome", "meals_fastfood", "sleep_hr")

# convert categorical variables into factors
df_adult[categorical_features] = lapply(df_adult[categorical_features], factor)

apply(df_adult, 2, function(x) sum(is.na(x))) # check how many missing data
sapply(df_adult, class) # check data classes


```

### Appendix II: Exploratory Data Analysis
#### Response Variable (Weight)
```{r fig.height=3, fig.width=10, fig.align="center", warning=FALSE}
library(ggplot2)
require(gridExtra)
transform1 = log(df_adult$weight)
transform2 = log(log(df_adult$weight))

# response variable distribution
plot1 = ggplot(df_adult, aes(weight)) +
  geom_histogram() +
  ggtitle("BMI")
plot2 = ggplot(df_adult, aes(transform1)) +
  geom_histogram() +
  ggtitle("Log(BMI)")
plot3 = ggplot(df_adult, aes(transform2)) +
  geom_histogram() +
  ggtitle("Log(Log(BMI))")
grid.arrange(plot1, plot2, plot3, ncol=3)
```


#### Predictor Variables

##### Categorical Variables
```{r fig.height=20, fig.width=15, fig.align="center", warning=FALSE}
plot1 = ggplot(df_adult, aes(x=gender, y = log(weight))) + geom_boxplot() + 
  theme(axis.text.x = element_text(angle=90)) + 
  labs(x="Gender", y="log(weight)") + 
  ggtitle("Weight by Gender")
plot2 = ggplot(df_adult, aes(x=race, y = log(weight))) + geom_boxplot() + 
  theme(axis.text.x = element_text(angle=90)) + 
  labs(x="Race", y="log(weight)") + 
  ggtitle("Weight by Race")
plot3 = ggplot(df_adult, aes(x=edu, y = log(weight))) + geom_boxplot() + 
  theme(axis.text.x = element_text(angle=90)) + 
  labs(x="Education Level", y="log(weight)") + 
  ggtitle("Weight by Education Level")
plot4 = ggplot(df_adult, aes(x=marriage, y = log(weight))) + geom_boxplot() + 
  theme(axis.text.x = element_text(angle=90)) + 
  labs(x="Marital Status", y="log(weight)") + 
  ggtitle("Weight by Marital Status")
plot5 = ggplot(df_adult, aes(x=famincome, y = log(weight))) + geom_boxplot() + 
  theme(axis.text.x = element_text(angle=90)) + 
  labs(x="Family Income", y="log(weight)") + 
  ggtitle("Weight by Family Income")
plot6 = ggplot(df_adult, aes(x=alcohol12yr, y = log(weight))) + geom_boxplot() + 
  theme(axis.text.x = element_text(angle=90)) + 
  labs(x="Alcohol Use", y="log(weight)") + 
  ggtitle("Weight by Alcohol Use")
plot7 = ggplot(df_adult, aes(x=milk, y = log(weight))) + geom_boxplot() + 
  theme(axis.text.x = element_text(angle=90)) + 
  labs(x="Milk Consumption", y="log(weight)") + 
  ggtitle("Weight by Milk Consumption")
plot8 = ggplot(df_adult, aes(x=depressed, y = log(weight))) + geom_boxplot() + 
  theme(axis.text.x = element_text(angle=90)) + 
  labs(x="Depressed", y="log(weight)") + 
  ggtitle("Weight by Despressed")
plot9 = ggplot(df_adult, aes(x=sleep_trouble, y = log(weight))) + geom_boxplot() + 
  theme(axis.text.x = element_text(angle=90)) + 
  labs(x="Alcohol Use", y="log(weight)") + 
  ggtitle("Weight by Sleep Trouble")
plot10 = ggplot(df_adult, aes(x=smoke, y = log(weight))) + geom_boxplot() + 
  theme(axis.text.x = element_text(angle=90)) + 
  labs(x="Smoke", y="log(weight)") + 
  ggtitle("Weight by Smoke")
plot11 = ggplot(df_adult, aes(x=tv_hrs, y = log(weight))) + geom_boxplot() + 
  theme(axis.text.x = element_text(angle=90)) + 
  labs(x="TV Consumption", y="log(weight)") + 
  ggtitle("Weight by TV Consumption")
grid.arrange(plot1, plot2, plot3, plot4, plot5, plot6, plot7, plot8, plot9, plot10, plot11, ncol=3)
```

##### Numeric Variables
Distribtuion of numeric variables
```{r fig.height=10, fig.width=10, fig.align="center", warning=FALSE}
# numeric predictor variable distributions
plot1 = ggplot(df_adult, aes(age)) + geom_histogram() +
  ggtitle("Age")
plot2 = ggplot(df_adult, aes(height)) + geom_histogram() +
  ggtitle("Height")
plot3 = ggplot(df_adult, aes(famsize)) + geom_histogram() +
  ggtitle("Family Size")
plot4 = ggplot(df_adult, aes(alcoholfrq)) + geom_histogram() +
  ggtitle("Frequency of Alcohol Consumption")
plot5 = ggplot(df_adult, aes(grocery)) + geom_histogram() +
  ggtitle("Money Spent on Grocery")
plot6 = ggplot(df_adult, aes(eatout)) + geom_histogram() +
  ggtitle("Money Spent Eating Out")
plot7 = ggplot(df_adult, aes(delivery)) + geom_histogram() +
  ggtitle("Money Spent on Delivered Food")
plot8 = ggplot(df_adult, aes(meals_nothome)) + geom_histogram() +
  ggtitle("Number of Meals Not Home Prepared")
plot9 = ggplot(df_adult, aes(meals_fastfood)) + geom_histogram() +
  ggtitle("Number of Fast Food Meals")
plot10 = ggplot(df_adult, aes(sleep_hr)) + geom_histogram() +
  ggtitle("Hours of Sleep")
grid.arrange(plot1, plot2, plot3, plot4, plot5, plot6, plot7, plot8, plot9, plot10, ncol=3)
```                    
          
```{r fig.height=20, fig.width=9, fig.align="center", warning=FALSE}
# transformations on numeric predictor variables
plot1 = ggplot(df_adult, aes(alcoholfrq)) + geom_histogram() +
  ggtitle("Frequency of Alcohol Consumption")
plot2 = ggplot(df_adult, aes(log(alcoholfrq)+1)) + geom_histogram() +
  ggtitle("Log Frequency of Alcohol Consumption")

plot3 = ggplot(df_adult, aes(grocery)) + geom_histogram() +
  ggtitle("Money Spent on Grocery")
plot4 = ggplot(df_adult, aes(log(grocery)+1)) + geom_histogram() +
  ggtitle("Log Money Spent on Grocery")

plot5 = ggplot(df_adult, aes(eatout)) + geom_histogram() +
  ggtitle("Money Spent on Eating Out")
plot6 = ggplot(df_adult, aes(log(eatout)+1)) + geom_histogram() +
  ggtitle("Log Money Spent on Eating Out")

plot7 = ggplot(df_adult, aes(delivery)) + geom_histogram() +
  ggtitle("Money Spent on Delivered Food")
plot8 = ggplot(df_adult, aes(log(delivery)+1)) + geom_histogram() +
  ggtitle("Log Money Spent on Delivered Food")

plot9 = ggplot(df_adult, aes(meals_nothome)) + geom_histogram() +
  ggtitle("Number of Meals Not Home Prepared")
plot10 = ggplot(df_adult, aes(log(meals_nothome)+1)) + geom_histogram() +
  ggtitle("Log Number of Meals Not Home Prepared")

plot11 = ggplot(df_adult, aes(meals_fastfood)) + geom_histogram() +
  ggtitle("Number of Fast Food Meals")
plot12 = ggplot(df_adult, aes(log(meals_fastfood)+1)) + geom_histogram() +
  ggtitle("Log Number of Fast Food Meals")
grid.arrange(plot1, plot2, plot3, plot4, plot5, plot6, plot7, plot8, plot9, plot10, plot11, plot12, ncol=2)
```  

Response vs. numeric distribution
```{r fig.height=10, fig.width=10, fig.align="center", warning=FALSE}
plot1 = ggplot(df_adult, aes(x=age, y=log(weight))) + geom_point() + 
  ggtitle("Age")
plot2 = ggplot(df_adult, aes(x=height, y=log(weight))) + geom_point() + 
  ggtitle("Height")
plot3 = ggplot(df_adult, aes(x=famsize, y=log(weight))) + geom_point() + 
  ggtitle("Family Size")
plot4 = ggplot(df_adult, aes(x=alcoholfrq, y=log(weight))) + geom_point() + 
  ggtitle("Frequency of Alcohol Consumption")
plot5 = ggplot(df_adult, aes(x=grocery, y=log(weight))) + geom_point() + 
  ggtitle("Money Spent on Grocery")
plot6 =ggplot(df_adult, aes(x=eatout, y=log(weight))) + geom_point() + 
  ggtitle("Money Spent Eating Out")
plot7 = ggplot(df_adult, aes(x=delivery, y=log(weight))) + geom_point() + 
  ggtitle("Money Spent on Delivered Food")
plot8 =ggplot(df_adult, aes(x=meals_nothome, y=log(weight))) + geom_point() + 
  ggtitle("Number of Meals Not Home Prepared")
plot9 = ggplot(df_adult, aes(x=meals_fastfood, y=log(weight))) + geom_point() + 
  ggtitle("Number of Fast Food Meals")
plot10 = ggplot(df_adult, aes(x=sleep_hr, y=log(weight))) + geom_point() + 
  ggtitle("Hours of Sleep")
grid.arrange(plot1, plot2, plot3, plot4, plot5, plot6, plot7, plot8, plot9, plot10, ncol=3)
```          


```{r fig.height=10, fig.width=10, fig.align="center", warning=FALSE}
plot1 = ggplot(df_adult, aes(x=age, y=log(weight), colour=gender)) + geom_point() + 
  ggtitle("Age")
plot2 = ggplot(df_adult, aes(x=height, y=log(weight), colour=gender)) + geom_point() + 
  ggtitle("Height")
plot3 = ggplot(df_adult, aes(x=famsize, y=log(weight), colour=gender)) + geom_point() + 
  ggtitle("Family Size")
plot4 = ggplot(df_adult, aes(x=log(alcoholfrq), y=log(weight), colour=gender)) + geom_point() + 
  ggtitle("Frequency of Alcohol Consumption")
plot5 = ggplot(df_adult, aes(x=grocery, y=log(weight), colour=gender)) + geom_point() + 
  ggtitle("Money Spent on Grocery")
plot6 =ggplot(df_adult, aes(x=eatout, y=log(weight), colour=gender)) + geom_point() + 
  ggtitle("Money Spent Eating Out")
plot7 = ggplot(df_adult, aes(x=delivery, y=log(weight), colour=gender)) + geom_point() + 
  ggtitle("Money Spent on Delivered Food")
plot8 =ggplot(df_adult, aes(x=meals_nothome, y=log(weight), colour=gender)) + geom_point() + 
  ggtitle("Number of Meals Not Home Prepared")
plot9 = ggplot(df_adult, aes(x=meals_fastfood, y=log(weight), colour=gender)) + geom_point() + 
  ggtitle("Number of Fast Food Meals")
plot10 = ggplot(df_adult, aes(x=sleep_hr, y=log(weight), colour=gender)) + geom_point() + 
  ggtitle("Hours of Sleep")
grid.arrange(plot1, plot2, plot3, plot4, plot5, plot6, plot7, plot8, plot9, plot10, ncol=3)
```          


```{r data-transformation, warning=FALSE}
# response
df_adult$log.weight = log(df_adult$weight)

# numeric predictors
df_adult$log.alcoholfrq = log(df_adult$alcoholfrq + 1)
df_adult$log.grocery = log(df_adult$grocery + 1)
df_adult$log.eatout = log(df_adult$eatout + 1)
df_adult$log.delivery = log(df_adult$delivery + 1)
df_adult$log.nothome = log(df_adult$meals_nothome + 1)
df_adult$log.fastfood = log(df_adult$meals_fastfood + 1)
```



parallel corrd plot to figure out if interaction may be helpful
```{r fig.height=5, fig.width=10, fig.align="center", warning=FALSE}
# library(MASS)
# library(RColorBrewer)
# k <- adjustcolor(brewer.pal(5, "Set2")[df_adult$sleep_trouble], alpha=.6)
# predictor = c("sleep_hr", "age", "log.delivery", "meals_nothome", "meals_fastfood")
# parcoord(df_adult[,predictor], col=k)
```

